### 2022-11-27 dissertation/chapters/06-bayesian-hyper-heuristic.tex
🚧 (final-review): Chapter 6 WIP

The above quote was the inspiration for the development of a novel \acf{HH} that uses Bayesian probability concepts as a selection mechanism to drive the heuristic selection process. Thus far the reader has been presented with all of the necessary background information on \acp{ANN} (Chapter \ref{chap:anns}), low-level \index{heuristic}heuristics (Chapter \ref{chap:heuristics}), \acp{HH} (Chapter \ref{chap:hhs}) and lastly, probability theory (Chapter \ref{chap:probability}). These elements form the fundamental components of the proposed \Acf{BHH}. This chapter provides the details around the implementation of the \Ac{BHH} and explains how it used to train \acp{FFNN}. The remainder of the chapter is structured as follows:

@@

      \item \textbf{Section \ref{sec:bhh:heuristic_pool}} presents the \index{heuristic}\textit{heuristic pool}, a collection of low-level \index{heuristic}heuristics. Discussions follow on the importance of diversity amoungst heuristics in the \index{heuristic pool}heuristic pool.

@@

The following sections presents the reader with more detail on the implementation of the \acs{BHH}. Discussions lead with the \acs{BHH} architecture and general \acs{HH} framework in the following section.

@@

This section aims to present the reader with all the high level components in the architecture of the \Acs{BHH}. Burke et al.~\cite{ref:burke:2010} proposed an initial framework for \acp{HH} and Grobler~\cite{ref:grobler:2015} further proposed a framework for a heterogeneous meta-\ac{HH}. The aforementioned frameworks are adapted for the implementation of the \Acs{BHH}. An illustration of the high-level architecture of the \Acs{BHH} is given in Figure \ref{fig:bhh_architecture}.

@@

With reference to Figure \ref{fig:bhh_architecture}, the components are briefly given below in the order of information flow during the \acs{FFNN} training process steps.

@@

      \item \textbf{Heuristic-Entity Selections}: Every entity in the \index{entity pool}entity pool gets assigned a selected heuristic.

@@

Generally speaking, the \index{heuristic pool}heuristic pool is a collection of low-level heuristics under consideration by the \acs{BHH}. The \index{heuristic pool}heuristic pool contains the set of low-level \index{heuristic}heuristics that, together with their performance information, make up the \text{heuristic space}. Importantly, the \acs{BHH} searches in \index{heuristic}heuristic space. The \index{heuristic pool}heuristic pool is defined in terms of the diversity of heuristics, as well as the amount of heuristics in the \index{heuristic pool}heuristic pool. Each of these are discussed in the following sections.

@@

A \index{heuristic}heuristic's eligibility for inclusion in the \index{heuristic}heuristic pool, is determined by its ability to solve to underlying problem and its search behaviour and characteristics. The \index{heuristic pool}heuristic pool should contain a variety of different heuristics that have different search behaviours, so that a trade-off can be made by between exploration and exploitation of solutions, as is required during the training process. Heuristics that explore a lot, should be selected when exploration is needed, and heuristics that exploit a lot should be selected accordingly. The trade-off between exploration and exploitation is managed by the high-level heuristic. The \acs{BHH} will learn to select and apply the appropriate heuristic, at the appropriate points, throughout the training process. In doing so, a balance is achieved between exploration and exploitation.

@@

Another design decision in the heuristic pool is heuristic pool size. Since the \acs{BHH} is a \textit{learning} \acs{HH}, every heuristic that is included in the heuristic pool increases the requirement of more statistical evidence of heuristic performance. Not only does a large \index{heuristic pool}heuristic pool drastically complicate the learning process that is required by the \acs{BHH}, but it also drastically complicates the process of maintaining \textit{state}. State is the term used to describe data that is relevant to the training process, and is maintained between training steps. An example of \textit{local} state is an entity's position, while an example of \textit{global} state is the global best solution found thus far. The following section presents the concept of \textit{proxies} and their role in maintaining state.

@@

The concept of proxies arise from the sparsity of state as maintained by different \index{heuristic}heuristics. Since \index{heuristic}heuristics maintain (possibly) different states, there is an uncertainty of state transition when switching between \index{heuristic}heuristics. Consider an example where the \index{heuristic pool}heuristic pool consists of just two \index{heuristic}heuristics. One heuristic is a gradient-based heuristic that maintains momentum such as \acs{Adam}. The other is a \index{meta-heuristic}meta-heuristic that does not require a gradient, such as \acs{PSO}. Both these \index{heuristic}heuristics track different parameters in their state. For \acs{Adam}, the expected gradient mean and variance is maintained, while the \ac{PSO} maintains record of the gbest and pbest solutions of all of entities. A solution to state indifference, is to proxy heuristic state update operations. State is then maintained in two parts:

@@

Primary and proxied state parameters must be maintained together. Since entities represent candidate solutions, which are forms of state, entities are ideal candidates to store and maintain these state parameters. Entities are extended to include the primary state elements of \textit{all} the underlying low-level heuristics, as well as the candidate solution (weights) for the \acs{FFNN}. At each heuristic-entity application step, all state parameters, per entity, are updated either by primary method or by proxied method. The \acs{BHH} thus incorporates a mapping of proxied state update operation as given in the example in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example} below.

\begin{table}[htbp]
      \centering
      \caption{An example of mapping of proxied state update operation maintained by the \acs{BHH}.}
      \label{tab:bhh:heuristic_pool:proxy_mapping_example}%
      \par\bigskip
      \resizebox{0.5\textwidth}{!}{
            \begin{tabular}{ccccc}
                                                                  &   & \multicolumn{3}{c}{State Parameter}                                                                                                                                                                                      \\
                  \cmidrule{3-5}                                  &   & 1                                                                      & 2                                                                      & 3                                                                      \\
                  \cmidrule{3-5}    \multirow{3}[1]{*}{Heuristic} & A & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{B}  & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} \\
                                                                  & B & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{A}  \\
                                                                  & C & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{B}  & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{A}  \\
            \end{tabular}%
      }
\end{table}%

From the example given in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}, when \index{heuristic}heuristic A is selected, it will outsource state update operations from \index{heuristic}heuristic B, for state parameter 2. Heuristic B will outsource from \index{heuristic}heuristic A, for state parameter 3. Finally, \index{heuristic}heuristic C will outsource from \index{heuristic}heuristic A and B, for state parameters 2 and 3 respectively. In this way, all \index{heuristic}heuristics maintain all the state parameters.

Proxied state update operations is a simple concept in principle, but requires detailed decomposition of the \index{heuristic}heuristics included in the \index{heuristic pool}heuristic pool.  Overlapping and unique state parameters must be identified so that a proxy mapping, such as the one given in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}, can be constructed. A suggestion to simplify this process is to borrow concepts from the equations of motion from physics. These include \textit{position}, \textit{velocity}, \textit{acceleration} and \textit{momentum}. Expressing \index{heuristic}heuristic update steps according to these parameters drastically simplify the process. However, it is possible that \index{heuristic}heuristics implement unique state parameters that do not overlap, these have to be catered for in the proxy mapping.

State parameters and update operations should not be considered in isolation. An example is \textit{position} and \textit{velocity}. Consider, for example, \index{heuristic}heuristics such as \acf{DE} and \acfp{GA}. These \index{heuristic}heuristics recombine entities. In the aforementioned case, the concept of an equation of motion does not entirely make sense, since the displacement of its position is not a result of maintaining velocity or momentum, but rather by displacement through recombination. In this particular case, a solution is to apply the recombination operation to \textit{all} applicable state parameters as well, or to nullify the necessary state parameters. Unfortunately there is no general, automated solution to formulate these heuristic state overlaps. Each \index{heuristic}heuristic must be carefully considered. More details on the \index{entity pool}entity pool and entity state are presented in the following section.

@@

Entities represent candidate solutions to the model's trainable parameters (weights) and other \index{heuristic}heuristic-specific state parameters, as was discussed in Section \ref{sec:bhh:heuristic:proxies}. It can be said that entities implement \textit{local} state. It was mentioned that these entities can be treated as physical \textit{particles} in a hyper-dimensional search environment. Entities model concepts from physics. For example, the candidate solution is represented as the entity's position, and an entity's velocity and acceleration is analogous to the gradient and momentum of the entity respectfully. Examples of entity state parameters, as derived from various low-level \index{heuristic}heuristics, is given as follows.

\begin{itemize}
      \item \textbf{position}: General parameter that represents the actual candidate solution and is thus a primary state parameter for all \index{heuristic}heuristics.

@@

      \item \textbf{personal best position}: Parameter that tracks that best known position by the entity thus far. As required and maintained by \index{heuristic}heuristics such as \acs{PSO}.

      \item \textbf{personal best loss}: Parameter that tracks that best known loss by the entity thus far. As required and maintained by \index{heuristic}heuristics such as \acs{PSO}.

      \item \textbf{loss}: Parameter that tracks the loss as achieved by the the entity throughout training.
\end{itemize}

@@

The population state refers to a collection of parameters that are shared between the entities in the population. Population state is also referred to as \textit{global} state and represents the population's memory. The population state generally contains state parameters that are of importance to multiple \index{heuristic}heuristics, and usually tracks the state of the population and not individual \index{heuristic}heuristic. Some examples of population state that can arise from different \index{heuristic}heuristics is given below.

@@

      \item \textbf{rbest} and \textbf{rbest loss}: Refers to the best position and loss achieved by the population for the current replay buffer/window size.

@@

Entities in the entity pool are each assigned an associated \index{heuristic}heuristic. The \acs{BHH} selects from the \index{heuristic pool}heuristic pool a low-level \index{heuristic}heuristic to be applied to an individual entity. The outcome of this selection process is a mapping table that tracks which \index{heuristic}heuristic has been selected for which entity. The selection process is executed by the selection mechanism of the \acs{BHH}. These \index{heuristic}heuristic-entity combinations are applied to the underlying \acs{FFNN}. The \acs{BHH} tracks the performance of each of these combinations throughout the training process in a performance log. The performance log is presented in the following section.

@@

The \acs{BHH} incorporates a form of probabilistic modeling in its selection mechanism. Probability is calculated based on \index{heuristic}heuristic-entity performance over time. Evidence of \index{heuristic}heuristic-entity performance is thus required for the \acs{BHH} to learn. Since the performance log can become very big, only a sliding window of the performance history is maintained at each step in the learning process. The sliding window is also referred to as a \textit{replay} buffer, a term borrowed from the field of \acf{RL}. The replay window size is defined as a hyper-parameter of the \acs{BHH}. The performance log is then simply a table of events and metrics. The \acs{BHH} tracks the following metrics in the performance log.

@@

      \item \textbf{rbest loss}:  Keeps track of the \textit{replay} best loss. Thus, the best loss value achieved by all entity-\index{heuristic}heuristic combinations, over all mini-batch steps currently in the performance log/replay buffer.

@@

An example of the performance log implemented by the \acs{BHH} is given in Table \ref{tab:bhh:performance_log:performance_log_example} below.

\begin{table}[htbp]
      \centering
      \caption{An example of the performance log implemented by the \acs{BHH}, showing the first 5 entities, their allocated \index{heuristic}heuristics and their resulting performance measurements for step 1 of the training process.}
      \label{tab:bhh:performance_log:performance_log_example}%
      \par\bigskip
      \resizebox{\textwidth}{!}{
            \begin{tabular}{cccccccc}
                  \textbf{step} & \textbf{entity} & \textbf{heuristic} & \textbf{loss} & \textbf{ibest loss} & \textbf{pbest loss} & \textbf{rbest loss} & \textbf{gbest loss} \\
                  \midrule
                  1             & 1               & 1                  & 0.016444      & 0.016444            & 0.016444            & 0.016444            & 0.016444            \\
                  1             & 2               & 2                  & 0.337965      & 0.016444            & 0.337965            & 0.016444            & 0.016444            \\
                  1             & 3               & 1                  & 0.134781      & 0.016444            & 0.134781            & 0.016444            & 0.016444            \\
                  1             & 4               & 1                  & 0.998719      & 0.016444            & 0.998719            & 0.016444            & 0.016444            \\
                  1             & 5               & 3                  & 0.708702      & 0.016444            & 0.708702            & 0.016444            & 0.016444            \\
            \end{tabular}%
      }
\end{table}%

The performance log itself introduces a number of design considerations as well. The larger the performance log, the longer memory is retained throughout the training process. The performance log size, referred to as the \textit{replay window size}, controls the recency of evidence from which the \acs{BHH} must learn. If the replay window size is too small, it does not accumulate enough samples/evidence to statistically make accurate selections. If the replay window size is too big, a bias could exist for selecting \index{heuristic}heuristics that performed well in the past. Furthermore, it is not guaranteed that past performance is indicative of future performance during the training process. The performance log should be considered along with the selection of a credit assignment strategy. Credit assignment strategies are presented in the following section.

@@

The credit assignment strategy is a mechanism that assigns a discrete credit indicator to \index{heuristic}heuristics that perform well, based on their performance metrics such as loss. The credit assignment strategy implements the ``move acceptance'' process as proposed by~\citeauthor{ref:ozcan:2006}~\cite{ref:ozcan:2006,ref:ozcan:2008} and addresses the credit assignment problem as discussed by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}. A good credit assignment strategy will correctly allocate credit to the appropriate \index{heuristic}heuristic-entity combination. The credit assignment strategy to use is defined as a hyper-parameter. The \acs{BHH} implements the following credit assignment strategies to choose from.

@@

      \item \textbf{symmetric}: Credit is assigned to all entity-\index{heuristic}heuristic combinations, regardless of their performance. The symmetric credit assignment strategy does not randomly assign credit, but rather assigns credit to all events. In effect, no learning and performance-bias is achieved with the symmetric credit assignment strategy and thus, the symmetric credit assignment strategy is implemented as a basis for comparison. Comparing a credit assignment staretegy to the symmetric credit assignment indicates the ability of the other credit assignment strategy to have an effect on the learning process and performance of the \acs{BHH}.
\end{itemize}

The implementation of a credit assignment strategy is simply a function that translates the performance log from Table \ref{tab:bhh:performance_log:performance_log_example}, into Table \ref{tab:bhh:credit_assignment_strategy:credit_assignment_example}.

\begin{table}[htbp]
      \centering
      \caption{Credit assignment strategy output table showing \textit{ibest} credit assignment for the first 5 entities and their selected heuristics, for step 1 of the training process.}
      \label{tab:bhh:credit_assignment_strategy:credit_assignment_example}%
      \par\bigskip
      \resizebox{0.4\textwidth}{!}{
            \begin{tabular}{cccc}
                  \textbf{step} & \textbf{entity} & \textbf{heuristic} & \textbf{credit} \\
                  \midrule
                  1             & 1               & 1                  & true            \\
                  1             & 2               & 2                  & false           \\
                  1             & 3               & 1                  & false           \\
                  1             & 4               & 1                  & false           \\
                  1             & 5               & 3                  & false           \\
            \end{tabular}%
      }
\end{table}%

Credit is used by the selection mechanism of the \acs{BHH}. The selection mechanism is presented in the following section.

@@

This section aims to provide the reader with the detail around the selection mechanism as it is implemented by the \acs{BHH}. It is mentioned throughout this document that the \acs{BHH} implements a predictive model based on Bayesian probabilistic concepts. In probabilistic modeling it is necessary to identify all the random events that can occur.

@@

Observation of random events is treated as evidence. In the context of a Bayesian approach, this evidence is used to update prior beliefs. The \acs{BHH} distinguishes between the following events.

\begin{itemize}
      \item \textbf{$H$}: The event of observing \textit{heuristics}.
      \item \textbf{$E$}: The event of observing \textit{entities}.
      \item \textbf{$C$}: The event of observing \textit{credit assignments} that indicate that the credit assignment \textit{performance criteria} is met.
\end{itemize}

@@

The dependence between random events can have an impact on the probabilistic model that is implemented. For simplicity, the \acs{BHH} assumes independence between events, although the event $C$ is clearly dependent on the occurrence of $H$ and $E$. Furthermore, the \acs{BHH} assumes independence between steps  and treats each training step as if training has restarted. The \acs{BHH} implements a form of Naïve Bayes classifier, where the appropriate heuristic to assign to an entity is a classification problem. \citeauthor{ref:domingos:1997}~\cite{ref:domingos:1997} mentions that although Bayesian classifiers’ probability estimates are only optimal under quadratic loss if the independence assumption holds, the classifier itself can still be optimal under zero-one loss (misclassification rate), even when this assumption is violated by a wide margin. This means that independence can be assumed when the probabilistic model is used as a classifier.

\index{Bayes' Theorem}
\subsection{Bayes' Theorem}\label{sec:bhh:selection_mechanism:bayes_theorem}

Chapter \ref{chap:probability} presented \index{Bayes' Theorem}Bayes' Theorem, but for convenience, it is given again in Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem} below.

@@

Since the \acs{BHH} implements a Bayesian classifier, the \acs{BHH} is not concerned with actual probabilities. Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem} can thus be evaluated for proportionality. The resulting proportionality is expressed in Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem_prop_to} below.

@@

The following section discusses the default values for proxied update step operations.

>>>>

The above quote was the inspiration for the development of a novel \acf{HH} that uses Bayesian probability concepts as a selection mechanism to drive the heuristic selection process. Thus far the reader has been presented with all of the necessary background information on \acp{ANN} (Chapter \ref{chap:anns}), low-level \index{heuristic}heuristics (Chapter \ref{chap:heuristics}), \acp{HH} (Chapter \ref{chap:hhs}) and lastly, probability theory (Chapter \ref{chap:probability}). These elements form the fundamental components of the proposed \Acf{BHH}. This chapter provides the details around the implementation of the \Ac{BHH} and explains how it is used to train \acp{FFNN}. The remainder of the chapter is structured as follows:

@@

      \item \textbf{Section \ref{sec:bhh:heuristic_pool}} presents the \index{heuristic}\textit{heuristic pool}, a collection of low-level \index{heuristic}heuristics. Discussions follow on the importance of diversity amongst heuristics in the \index{heuristic pool}heuristic pool.

@@

@@

This section aims to present the reader with all the high level components in the architecture of the \Acs{BHH}. Burke et al.~\cite{ref:burke:2010} propose an initial framework for \acp{HH} and Grobler~\cite{ref:grobler:2015} further proposed a framework for a heterogeneous meta-\ac{HH}. The aforementioned frameworks are adapted for the implementation of the \Acs{BHH}. An illustration of the high-level architecture of the \Acs{BHH} is given in Figure \ref{fig:bhh_architecture}.

@@

With reference to Figure \ref{fig:bhh_architecture}, the components are briefly given below in the order of information flow during the \acs{FFNN} training process steps:

@@

      \item \textbf{Heuristic-Entity Selections}: Every entity in the \index{entity pool}entity pool is assigned a selected heuristic.

@@

Generally speaking, the \index{heuristic pool}heuristic pool is a collection of low-level heuristics under consideration by the \acs{BHH}. The \index{heuristic pool}heuristic pool contains the set of low-level \index{heuristic}heuristics that, together with their performance information, make up the \text{heuristic space}. Importantly, the \acs{BHH} searches in \index{heuristic}heuristic space. The \index{heuristic pool}heuristic pool is defined in terms of the diversity of heuristics, as well as the number of heuristics in the \index{heuristic pool}heuristic pool. Each of these are discussed in the following sections.

@@

A \index{heuristic}heuristic's eligibility for inclusion in the \index{heuristic}heuristic pool is determined by its ability to solve the underlying problem and its search behaviour and characteristics. The \index{heuristic pool}heuristic pool should contain a variety of different heuristics that have different search behaviours, so that a trade-off can be made between exploration and exploitation of solutions, as is required during the training process. Heuristics that explore a lot should be selected when exploration is needed, and heuristics that exploit a lot should be selected accordingly. The trade-off between exploration and exploitation is managed by the high-level heuristic. The \acs{BHH} learns to select and apply the appropriate heuristic at the appropriate time, throughout the training process. In doing so, a balance is achieved between exploration and exploitation.

@@

Another design decision with respect to the heuristic pool is heuristic pool size. Since the \acs{BHH} is a \textit{learning} \acs{HH}, every heuristic that is included in the heuristic pool increases the requirement of more statistical evidence of heuristic performance. Not only does a large \index{heuristic pool}heuristic pool drastically complicate the learning process that is required by the \acs{BHH}, but it also drastically complicates the process of maintaining \textit{state}. State is the term used to describe data that is relevant to the training process, and is maintained between training steps. An example of \textit{local} state is an entity's position, while an example of \textit{global} state is the global best solution found thus far.

@@

The concept of proxies arise from the sparsity of state as maintained by different \index{heuristic}heuristics. Since \index{heuristic}heuristics maintain (possibly) different states, there is an uncertainty of state transition when switching between \index{heuristic}heuristics. Consider an example where the \index{heuristic pool}heuristic pool consists of just two \index{heuristic}heuristics. One heuristic is a gradient-based heuristic that maintains momentum, such as \acs{Adam}~\cite{ref:kingma:2014}. The other is a \index{meta-heuristic}meta-heuristic that does not require a gradient, such as \acs{PSO}~\cite{ref:shi:1998}. Both these \index{heuristic}heuristics track different parameters in their state. For \acs{Adam}, the expected gradient mean and variance is maintained, while the \ac{PSO} maintains record of the gbest and pbest solutions of all of entities. A solution to state indifference is to proxy heuristic state update operations. State is then maintained in two parts:

@@

Primary and proxied state parameters must be maintained together. Since entities represent candidate solutions, which are forms of state, entities are ideal candidates to store and maintain these state parameters. Entities are extended to include the primary state elements of \textit{all} the underlying low-level heuristics, as well as the candidate solution (weights) for the \acs{FFNN}. At each heuristic-entity application step, all state parameters, per entity, are updated either by primary method or by proxied method. The \acs{BHH} thus incorporates a mapping of proxied state update operation as given in the example in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}.

\begin{table}[htbp]
      \centering
      \caption{An example of mapping of proxied state update operation maintained by the \acs{BHH}.}
      \label{tab:bhh:heuristic_pool:proxy_mapping_example}%
      \par\bigskip
      \resizebox{0.38\textwidth}{!}{
            \begin{tabular}{ccccc}
                                                                  &   & \multicolumn{3}{c}{State Parameter}                                                                                                                                                                                      \\
                  \cmidrule{3-5}                                  &   & 1                                                                      & 2                                                                      & 3                                                                      \\
                  \cmidrule{3-5}    \multirow{3}[1]{*}{Heuristic} & A & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{B}  & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} \\
                                                                  & B & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{A}  \\
                                                                  & C & \cellcolor[rgb]{ .776,  .937,  .808}\textcolor[rgb]{ 0,  .38,  0}{n/a} & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{B}  & \cellcolor[rgb]{ 1,  .922,  .612}\textcolor[rgb]{ .612,  .341,  0}{A}  \\
            \end{tabular}%
      }
\end{table}%

From the example given in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}, when \index{heuristic}heuristic A is selected, it will outsource state update operations from \index{heuristic}heuristic B for state parameter 2. Heuristic B will outsource from \index{heuristic}heuristic A for state parameter 3. Finally, \index{heuristic}heuristic C will outsource from \index{heuristic}heuristic A and B for state parameters 2 and 3 respectively. In this way, all \index{heuristic}heuristics maintain all the state parameters.

Proxied state update operations is a simple concept in principle, but requires detailed decomposition of the \index{heuristic}heuristics included in the \index{heuristic pool}heuristic pool. Overlapping and unique state parameters must be identified so that a proxy mapping, such as the one given in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}, can be constructed. A suggestion to simplify this process is to borrow concepts from the equations of motion from physics. These include \textit{position}, \textit{velocity}, \textit{acceleration}, and \textit{momentum}. Expressing \index{heuristic}heuristic update steps according to these parameters drastically simplify the process. However, it is possible that \index{heuristic}heuristics implement unique state parameters that do not overlap. These have to be catered for in the proxy mapping.

State parameters and update operations should not be considered in isolation. An example is \textit{position} and \textit{velocity}. Consider, for example, \index{heuristic}heuristics such as \acs{DE}~\cite{ref:price:2006} and \acp{GA}~\cite{ref:holland:1992}. These \index{heuristic}heuristics recombine entities. In the aforementioned case, the concept of an equation of motion does not entirely make sense, since the displacement of its position is not a result of maintaining velocity or momentum, but rather by displacement through recombination. In this particular case, a solution is to apply the recombination operation to \textit{all} applicable state parameters as well, or to nullify the necessary state parameters. Unfortunately, there is no general, automated solution to formulate these heuristic state overlaps. Each \index{heuristic}heuristic must be carefully considered.

@@

Entities represent candidate solutions to the model's trainable parameters (weights) and other \index{heuristic}heuristic-specific state parameters, as was discussed in Section \ref{sec:bhh:heuristic:proxies}. It can be said that entities implement \textit{local} state. It was mentioned that these entities can be treated as physical \textit{particles} in a hyper-dimensional search environment. Entities model concepts from physics. For example, the candidate solution is represented as the entity's position, and an entity's velocity and acceleration is analogous to the gradient and momentum of the entity respectfully. Examples of entity state parameters, as derived from various low-level \index{heuristic}heuristics, is given as follows:

\begin{itemize}
      \item \textbf{position}: A general parameter that represents the actual candidate solution and is thus a primary state parameter for all \index{heuristic}heuristics.

@@

      \item \textbf{personal best position}: A parameter that tracks that best known position by the entity thus far, as required and maintained by \index{heuristic}heuristics such as \acs{PSO}.

      \item \textbf{personal best loss}: A parameter that tracks that best known loss by the entity thus far, as required and maintained by \index{heuristic}heuristics such as \acs{PSO}.

      \item \textbf{loss}: A parameter that tracks the loss as achieved by the the entity throughout training.
\end{itemize}

@@

The population state refers to a collection of parameters that are shared between the entities in the population. Population state is also referred to as \textit{global} state and represents the population's memory. The population state generally contains state parameters that are of importance to multiple \index{heuristic}heuristics, and usually tracks the state of the population and not individual \index{heuristic}heuristic. Some examples of population state that can arise from different \index{heuristic}heuristics is given below:

@@

      \item \textbf{rbest} and \textbf{rbest loss}: Refers to the best position and loss achieved by the population for a number of steps, referred to as the replay window size. The replay window size defines a window of memory for tracking historical performance data of \index{heuristic}heuristics.

@@

Entities in the entity pool are each assigned an associated \index{heuristic}heuristic. The \acs{BHH} selects from the \index{heuristic pool}heuristic pool a low-level \index{heuristic}heuristic to be applied to an individual entity. The outcome of this selection process is a mapping table that tracks which \index{heuristic}heuristic has been selected for which entity. The selection process is executed by the selection mechanism of the \acs{BHH}. These \index{heuristic}heuristic-entity combinations are applied to the underlying \acs{FFNN}. The \acs{BHH} tracks the performance of each of these combinations throughout the training process in a performance log.

@@

The \acs{BHH} incorporates a form of probabilistic modeling in its selection mechanism. Probability is calculated based on \index{heuristic}heuristic-entity performance over time. Evidence of \index{heuristic}heuristic-entity performance is thus required for the \acs{BHH} to learn. Since the performance log can become very big, only a sliding window of the performance history is maintained at each step in the learning process. The sliding window is also referred to as a \textit{replay} buffer, a term borrowed from the field of \acf{RL}. The replay window size is defined as a hyper-parameter of the \acs{BHH}. The performance log is then simply a table of events and metrics. The \acs{BHH} tracks the following metrics in the performance log:

@@

      \item \textbf{rbest loss}:  Keeps track of the \textit{replay} best loss. Thus, the best loss value achieved by all entity-\index{heuristic}heuristic combinations, over all mini-batch steps currently in the performance log/replay window.

@@

An example of the performance log implemented by the \acs{BHH} is given in Table \ref{tab:bhh:performance_log:performance_log_example}.

\begin{table}[htbp]
      \centering
      \caption{An example of the performance log implemented by the \acs{BHH}, showing the first five entities, their allocated \index{heuristic}heuristics and their resulting performance measurements for step 1 of the training process.}
      \label{tab:bhh:performance_log:performance_log_example}%
      \par\bigskip
      \resizebox{\textwidth}{!}{
            \begin{tabular}{cccccccc}
                  \textbf{step} & \textbf{entity} & \textbf{heuristic} & \textbf{loss} & \textbf{ibest loss} & \textbf{pbest loss} & \textbf{rbest loss} & \textbf{gbest loss} \\
                  \midrule
                  1             & 1               & 1                  & 0.016444      & 0.016444            & 0.016444            & 0.016444            & 0.016444            \\
                  1             & 2               & 2                  & 0.337965      & 0.016444            & 0.337965            & 0.016444            & 0.016444            \\
                  1             & 3               & 1                  & 0.134781      & 0.016444            & 0.134781            & 0.016444            & 0.016444            \\
                  1             & 4               & 1                  & 0.998719      & 0.016444            & 0.998719            & 0.016444            & 0.016444            \\
                  1             & 5               & 3                  & 0.708702      & 0.016444            & 0.708702            & 0.016444            & 0.016444            \\
            \end{tabular}%
      }
\end{table}%

The performance log itself introduces a number of design considerations as well. The larger the performance log, the longer memory is retained throughout the training process. The performance log size, referred to as the \textit{replay window size}, controls the recency of evidence from which the \acs{BHH} must learn. If the replay window size is too small, it does not accumulate enough samples/evidence to statistically make accurate selections. If the replay window size is too big, a bias could exist for selecting \index{heuristic}heuristics that performed well in the past. Furthermore, it is not guaranteed that past performance is indicative of future performance during the training process. The performance log should be considered along with the selection of a credit assignment strategy.

@@

The credit assignment strategy is a mechanism that assigns a discrete credit indicator to \index{heuristic}heuristics that perform well, based on their performance metrics such as loss. The credit assignment strategy implements the ``move acceptance'' process as proposed by~\citeauthor{ref:ozcan:2006}~\cite{ref:ozcan:2006,ref:ozcan:2008} and addresses the credit assignment problem as discussed by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}. A good credit assignment strategy will correctly allocate credit to the appropriate \index{heuristic}heuristic-entity combination. The credit assignment strategy to use is defined as a hyper-parameter. The \acs{BHH} implements the following credit assignment strategies to choose from:

@@

      \item \textbf{symmetric}: Credit is assigned to all entity-\index{heuristic}heuristic combinations, regardless of their performance. The symmetric credit assignment strategy does not randomly assign credit, but rather assigns credit to all events. In effect, no learning and performance-bias is achieved with the symmetric credit assignment strategy and thus the symmetric credit assignment strategy is implemented as a basis for comparison. Comparing a credit assignment staretegy to the symmetric credit assignment indicates the ability of the other credit assignment strategy to have an effect on the learning process and performance of the \acs{BHH}.
\end{itemize}

The implementation of a credit assignment strategy is simply a function that translates the performance log from Table \ref{tab:bhh:performance_log:performance_log_example} into Table \ref{tab:bhh:credit_assignment_strategy:credit_assignment_example}.

\begin{table}[htbp]
      \centering
      \caption{Credit assignment strategy output table showing \textit{ibest} credit assignment for the first five entities and their selected heuristics for step 1 of the training process.}
      \label{tab:bhh:credit_assignment_strategy:credit_assignment_example}%
      \par\bigskip
      \resizebox{0.4\textwidth}{!}{
            \begin{tabular}{cccc}
                  \textbf{step} & \textbf{entity} & \textbf{heuristic} & \textbf{credit} \\
                  \midrule
                  1             & 1               & 1                  & true            \\
                  1             & 2               & 2                  & false           \\
                  1             & 3               & 1                  & false           \\
                  1             & 4               & 1                  & false           \\
                  1             & 5               & 3                  & false           \\
            \end{tabular}%
      }
\end{table}%

Credit is used by the selection mechanism of the \acs{BHH}.

@@

This section provides the detail around the selection mechanism as it is implemented by the \acs{BHH}. Detail is provided around random events as observed by the \acs{BHH}. An argument for independence between random events is made. \index{Bayes' theorem}Bayes' theorem is briefly reviewed in the context of the \acs{BHH} implementation. It is shown how the \acs{BHH} implements a probabilistic, predictive model based on the fundamentals of the Na\"{\i}ve Bayes algorithm. Brief discussions follow on numerical stability and mode collapse.

@@

Observation of random events is treated as evidence. In the context of a Bayesian approach, this evidence is used to update prior beliefs. The \acs{BHH} distinguishes between the following events:

\begin{itemize}
      \item \textbf{$H$}: The event of observing \textit{heuristics}.
      \item \textbf{$E$}: The event of observing \textit{entities}.
      \item \textbf{$C$}: The event of observing \textit{credit assignments} that indicate that the credit assignment \textit{performance criteria} are met.
\end{itemize}

@@

The dependence between random events can have an impact on the probabilistic model that is implemented. For simplicity, the \acs{BHH} assumes independence between events, although the event $C$ is clearly dependent on the occurrence of $H$ and $E$. Furthermore, the \acs{BHH} assumes independence between steps  and treats each training step as if training has restarted. The \acs{BHH} implements a form of Naïve Bayes classifier, where the appropriate heuristic to assign to an entity is a classification problem. \citeauthor{ref:domingos:1997}~\cite{ref:domingos:1997} mention that although the probability estimates of Bayesian classifiers are only optimal under quadratic loss if the independence assumption holds, the classifier itself can still be optimal under zero-one loss (misclassification rate), even when this assumption is violated by a wide margin. This means that independence can be assumed when the probabilistic model is used as a classifier.

\index{Bayes' theorem}
\subsection{Bayes' Theorem}\label{sec:bhh:selection_mechanism:bayes_theorem}

Chapter \ref{chap:probability} presented \index{Bayes' theorem}Bayes' theorem, but for convenience, it is given below:

@@

Since the \acs{BHH} implements a Bayesian classifier, the \acs{BHH} is not concerned with actual probabilities. Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem} can thus be evaluated for proportionality. The resulting proportionality is expressed as

@@
